---
output: pdf_document
---
## Analyses of tooth growth
### Author: "G ARGENTON"

## Overview
Analyses of tooth growth as part of "Course project" of Statistical Inference course of John Hopkins university in Coursera

### Exploratory data analysis
Let's load the ToothGrowth data and perform some basic exploratory data analyses

```{r}
data(ToothGrowth)
summary(ToothGrowth)
```

### Summary of the data 
Let's analyse the average thooth growth by sub groups of specific dose and supp  
Let's plot the data to see patterns
```{r}
ToothGrowthByDoseSupp <- aggregate(len ~ dose + supp,data=ToothGrowth,mean)
ToothGrowthByDoseSupp

library(ggplot2)
g <- ggplot(data = ToothGrowth, aes(x = dose, y = len, fill = supp ))
g <- g + geom_point(size =10, pch = 21, alpha = .5)
g
```
  
It seems that tooth growth is increasing with dosis, and that OJ is more efficient than VC.

### Comparaison on tooth growth by supp and dose. 
For supp VC, let's compare the group with 0.5 dosis and the group with 2 dosis
```{r}
# subseting data into two groups g1 & g2
g1 <- ToothGrowth[ToothGrowth$supp == "VC" & ToothGrowth$dose == 0.5,]$len
g2 <- ToothGrowth[ToothGrowth$supp == "VC" & ToothGrowth$dose == 2,]$len
print(paste("Average tooth growth supp VC dose 0.5 :", mean(g1),", dose 2 :",mean(g2)))
print(paste("Average increase of tooth growth using VC increasing dosis from 0.5 to 2 :",mean(g2)-mean(g1)))
# Let's calculate a 95 % student's t confidence interval for two independant groups
# we assume constant variance
sd1 <- sd(g1); sd2 <- sd(g2)
pv <- (9*sd1^2+9*sd2^2)/18
semd <- sqrt(pv)*sqrt(1/10+1/10)
round(mean(g2)-mean(g1) + c(-1,1)*qt(0.975,18)*semd,2)
```

Let's test the hypothesis of more efficiency with supp OJ than VC
```{r}
# subseting data into two groups VC & OJ
VC <- ToothGrowth[ToothGrowth$supp == "VC" ,]$len
OJ <- ToothGrowth[ToothGrowth$supp == "OJ", ]$len
print(paste("Average tooth growth supp VC :", round(mean(VC),2),"supp OJ:", round(mean(OJ),2)))
print(paste("Average increase of tooth growth using OJ vs VC :",mean(OJ)-mean(VC)))
# Let's test the hypothesis of one supp being more efficient (two sided test) with alpha = 5% 
# the two groups are independant, and we assume constant variance  
t.test(OJ, VC, paired = FALSE, var.equal = TRUE)
qt(p = 0.975,df = 58)
```


### Conclusions and the assumptions needed for conclusions. 
Tooth lengh is higher with VC and doses = 2 than with VC and dose = 0.5 : with 95% confidence, increase is between 14.49 and 21.83

We can not conclude that "one supp is more efficient than the other" with alpha = 5% : 
we fail to reject the null hypothesis (O within 95% confidence interval, or statistic 1.9153 < 2.0017 t quantile, or p-value = 0.06 > 0.05)

NB : We have assumed that distribution is gaussian or at least symetric and mound shaped so that t tests and Student's t confidence interval applies, and data are resonably iid sample of the polupation. We also assume that variance is constant within the groups.

